---
title: "Probability Distributions in R"
author: "Serdar Korur"
date: "12/3/2019"
output: html_document
---

Univariate probability distributions:

1. Normal distribution
2. Binomial distribution
3. Multinomial distribution
3. Poissson distribution
4. Expontential distribution
5. Geometric distribution
6. Gamma distribution
3

Multivariate probability distributions:


# Multivariate distribution

Multivariate distributions is used  to describe probability distributions of more than one variable at the same time allowing us to see the relation between them. 

### When we need them?

In hypothesis testing, clustering, and likelihood calculation, you need
to calculate the density of a specified multivariate normal distribution

Student courses, does physics correlates with math results or art with english.

### Mean vector

Since there are more than one variable mean is not a single value but a vector with the length of number of variables

### variance-covariance matrix

**Variance** is the average of the squared differences from the Mean and
we use variance to describe the spread the a univariate distribution and variance-covariance matrix to describe the spread of multivariate data. 

**relation**
sd to variance
correlation to covariance

### Why we need a Variance-Covariance matrix

**Variance and covariance matrix** is used to store the spread parameters of a multivariate data. 
Variances of individual random variables appear at the diagonal and covariances appear in the off-diagonal elements. 

Covariance represents the product of the differences of two variables from their mean. It measures how two variables of a data move together. 

when we standardise the covariance of two variables by dividing with the product of their individual standard deviations we get the correlation value.

>>Formula..


var() is used to calculate variance co variance matrix.
cor() is used to calculate correlation matrix
corrplot()
### What is the difference between univariate data vs multivariate data?

Univariate distributions a
Univariate normal vs multivariate normal

### R functions for multivariate data

library(mvtnorm)
rmvnorm(n, mean , sigma)

rnorm rmvnorm rt rmvt
qnorm
dnorm
pnorm
## Definitions

### Variance-Covariance matrix
###  Correlation matrix

## Plotting multivariate variables
pairs()
splom()
ggpairs()
scatterplot3d()

$$\sum_{i=1}^n X_i$$
$$E[X{i}-\hat X][Y-\hat Y] $$

$$Var(X) =\sum_{}(X{i}-\hat X)^2/N$$
# Code for plotting bivariate densities
```{r}
# Create grid
d <- expand.grid(seq(-3, 6, length.out = 50 ), seq(-3,  6, length.out = 50))                   

# Calculate density on grid
dens1 <- dmvnorm(as.matrix(d), mean=c(1,2), sigma=matrix(c(1, .5, .5, 2), 2))

# Convert to matrix 
dens1 <-  matrix(dens1, nrow = 50 )

# Use perspective plot
persp(dens1, theta = 80, phi = 30, expand = 0.6, shade = 0.2, 
        col = "lightblue", xlab = "x", ylab = "y", zlab = "dens")
```


#### What is Kurtosis?

Kurtosis is a statistical measure that defines how heavily the tails of a distribution differ from the tails of a normal distribution. In other words, kurtosis identifies whether the tails of a given distribution contain extreme values.

# Cumulative Distribution


# Multivariate Statistical techniques (that assume multivariate normality)

* multivariate regressiin
* discriminant analysis
* model-based clustering
* principal component analysis (PCA)
* multivariate analysis of variance (MANOVA)

**qqnorm** is widely used to test univariate normality. ( Created by plotting theoretical quantiles to the sample quantiles)

Calculate univariate normality of each variable in the data
### Multuvariate normality tests

# newer version does not work.
install.packages("devtools")
library(devtools)
install_version('MVN', '4.0.2')

### qqnorm of all variables

uniPlot(diabetes[,1:9])

#### Check multivariate normality

mardiaTest(diabetes[, 1:8])

#### Multivariate normal distribtuions use cases in machine learning

https://brilliant.org/wiki/multivariate-normal-distribution/

### e.g Anomaly detection???

### Other types of multivariate distributions

Not all multivatiate processess follow multivariate normal

multivariate-t distribtuion
skewed-normal-multivariate
skewed-t-multivariate

mean: is the location parameter
variane: is the scale parameter

normal vs t compared

Distribution	Probability
Normal	0.05
t(df=1)	0.3
t(df=8)	0.0857
t(df=20)	0.0641
t(df=30)	0.0593

R Functions for multivariate t distributions (delta,sigma)
rmvt
qmvt
dmvt
pmvt

## In the stock market, individual stocks are modeled my univariate t distribution
due to the volatility of prices
if you want to model 3 stocks, you need a multivatiate t distribution

## Calculating the density of a multivariate distribution on a grid
```{r}
x <- seq(-3, 6, by = 1); y <- seq(-3, 6, by = 1)
d <- expand.grid(x = x, y = x)                   
del1 <- c(1, 2); sig1 <- matrix(c(1, .5, .5, 2), 2)
dens <- dmvt(as.matrix(d), delta = del1, sigma = sig1, df = 10, log = FALSE)
scatterplot3d(cbind(d, dens), type = "h", zlab = "density")
```

### Skewed multivatiate

Both normal and t distributions are designed to model symetric distr.b
but real data is often skewed.

# Functions for skewed normal distribution

SN library
dmsn
pmsn
rmsn
need to specify(xi, omega, alpha)
xi is the location parameter, mean
omega is the scale parameter, variance, covariance
alpha= is the skweness parameter

# functions for skewed-t-distribution

dmst
pmst
rmst
Need to specify xi, Omega, alpha, nu (degrees of freedom)

## plot skewed-t-dist

# Specify xi, Omega and alpha
# Specify xi, Omega and alpha
xi1 <- c(1, 2, -5) 
Omega1 <- matrix(c(1, 1, 0,
                   1, 2, 0,
                   0, 0, 5), 3, 3)
alpha1 <- c(4, 30, -5)

# Generate samples
skew.sample <- rmsn(n = 2000, xi = xi1, Omega = Omega1, alpha = alpha1)

Estimation of parameters from data
Need iterative algorithm to estimate the parameters of a skew-normal distribution

No explicit equation to calculate parameters
Several functions in sn package, including msn.mle() function

## PCA: Why we need principal component analysis

Large number of strongly correlated variables, this poses

Computational challenges
and results are hard to interpret

to avoid this ds often choose to work with relevant but uncorrelated variables. 
weighted sums of existing variables.

princomp() calculates, principal components in the data. 

# Calculate PCs
pca.state <- princomp(state.x77, cor =TRUE, scores = TRUE) 

# Plot the PCA object  
plot(pca.state) 

# Print the summary of the PCs
summary(pca.state) 

## Find the Principal compoments needed to explain a certain level of variation in the data
Cumulative proportion

# Variance explained
pc.var <- cars.pca$sdev^2

# Proportion of variation
pc.pvar <- pc.var / sum(pc.var)

# Cumulative proportion
plot(cumsum(pc.pvar), type = 'b')

# Draw screeplot
screeplot(pca.state, type = "l")
grid()

# Quickly visualize first 2 principal components with bi plot

Calculating, visualizing and intrepreting scores
biplot(cars.pca, col = c("steelblue", "white"), cex = c(0.8, 0.01))

Plotting the PC scores in two dimensions is one way to visualize high dimensional data.

# Using factoextra library for visualizing PCA components


The factoextra library provides a number of functions which make it easy to extract and visualize the output of many exploratory multivariate data analyses, including PCA. In this exercise, you will use several functions to visualize the scores and PC loadings of the state.x77 dataset.

_Use the output of princomp() function_

# Plot the scores
fviz_pca_ind(pca.state)

# Plot the PC loadings
fviz_pca_var(pca.state)

# Create a biplot
fviz_pca_biplot(pca.state)


## Multi-dimensional scaling

Another way of dimension reduction. to provide visual representation

cmdscale(distance matrix, k=max dimension, ...)

## Multidimensional scaling in more than two dimensions
cars.dist <- dist(mtcars)
cmds3 <- data.frame(cmdscale(cars.dist, k = 3))
scatterplot3d(cmds3, type = "h", pch = 19, lty.hplot = 2, color = mtcars$cyl)

## Parameter estimation for multivariate skew-normals

Unlike multivariate normal, where the parameters estimates can be obtained using the sample mean and sample variance-covariance matrix, the parameters of the skew-normal distribution need to be estimated by an iterative process. You can make use of existing functions in the sn packages to get the estimates.


_________________________________________________________________________
=========================================================================
_________________________________________________________________________
# Sources:

Variance-co variance matrix

# https://stattrek.com/matrix-algebra/covariance-matrix.aspx

R Markdown for formulas
https://www.statpower.net/Content/310/R%20Stuff/SampleMarkdown.html
