---
title: "binomial_extras"
author: "Serdar Korur"
date: "10/31/2019"
output: html_document
---

```{r}
horror <- horror_date_table %>% ggplot(aes(x=n)) + geom_histogram()
horror_date_table %>% ggplot(aes(x=n)) + geom_density()
horror <- horror_date_table %>% mutate(percent = 100*n/2782)

p <- rbinom(100000, 2782,1/30)
p <- data.frame(p)
p %>% ggplot(aes(x=p)) + geom_density()


# Create pdfs

# Theoretical

pd <- density(p$p)
pd
mode <- pd$x[which.max((pd$y))]
mode
ggplot(p, aes(x=p)) + geom_rug() + geom_density() + geom_vline(xintercept = mode, col = "red")

fun_args <- list(mean = mean(p$p), sd = sd(p$p))


ggplot(p, aes(x = p)) +
  geom_histogram(aes(y = ..density..)) +
  geom_density(col = "red") +
  stat_function(fun = dnorm, args = fun_args, col = "blue")

# Empirical



hpdf <- density(horror_date_table$n)
mode_h <- hpdf$x[which.max(hpdf$y)]
ggplot(horror_date_table, aes(x=n)) + geom_rug() + geom_density() + geom_vline(xintercept = mode_h, col = "red") 

fun_args <- list(mean=mean(horror_date_table$n, sd= sd(horror_date_table$n)))

ggplot(horror_date_table, aes(x = n)) +
  geom_histogram(aes(y = ..density..)) +
  geom_density(col = "red") +
  stat_function(fun = dnorm, args = fun_args, col = "blue")
                 
```

```{r}
# QQplot sometimes can be misquiding
# Shapiro wilk test will show if normality is indeed the case
Shapiro.test()
```

**The performance of a machine learning algorithm on a binary classification problem can be analyzed as a Bernoulli process, where the prediction by the model on an example from a test set is a Bernoulli trial (correct or incorrect).**

The Binomial distribution summarizes the number of successes k in a given number of Bernoulli trials n, with a given probability of success for each trial p.

We can demonstrate this with a Bernoulli process where the probability of success is 30% or P(x=1) = 0.3 and the total number of trials is 100 (k=100).

# This is also good.

# example of using the pmf for the binomial distribution
from scipy.stats import binom
# define the parameters of the distribution
p = 0.3
k = 100
# define the distribution
dist = binom(k, p)
# calculate the probability of n successes
for n in range(10, 110, 10):
	print('P of %d success: %.3f%%' % (n, dist.pmf(n)*100))